### 2023-11-10 ms.Rmd
An early complete draft.

---
title: 'On investigating heterogeneity in psychological phenomena'
shorttitle: Heterogeneity in psychology
leftheader: Heterogeneity in psychology
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: Tilburg School of Social and Behavioral Sciences, Tilburg University
    email: mjvuorre@uvt.nl
  - name: Matthew Kay
    affiliation: 2
  - name: Niall Bolger
    affiliation: 3
affiliation:
  - id: 1
    institution: Department of Social Psychology, Tilburg School of Social and Behavioral Sciences, Tilburg University
  - id: 2
    institution: Computer Science and Communication Studies, Northwestern University
  - id: 3
    institution: Department of Psychology, Columbia University
abstract: |
  Advances in statistics and data collection methods have brought variability in psychological phenomena to the forefront of theoretical development. Yet, established methods and practices for studying heterogeneity underestimate the true uncertainty in those estimates. In the current work, we aim to deliver methods and guidelines for more rigorous treatments of heterogeneity in psychological phenomena. We provide a walkthrough of methods for i. estimating, ii. summarising, and iii. describing heterogeneity. Although other methods for incorporating estimation uncertainty are available, the methods we propose---involving bayesian hierarchical models---are especially suitable for investigating heterogeneity because they naturally describe uncertainty in all model parameters. We hope that the methods we discuss would help researchers to widen their inferential focus from statistical averages to include statistical heterogeneity in psychological phenomena.
keywords: variation, heterogeneity, hierarchical models, mixed models, methodology
authornote: |
  \textbf{This working paper is not yet peer reviewed.}
wordcount: "5,837"
bibliography: references.bib
floatsintext: yes
numbersections: false
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: jou
header-includes:
  - |
    \usepackage{float}
    \setlength{\parskip}{0.5em}
output: 
  papaja::apa6_pdf:
    number_sections: true
    keep_tex: false
    highlight: kate
  papaja::apa6_docx: 
    number_sections: true
    keep_tex: false
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(janitor)
library(scales)
library(ellipse)
library(knitr)
library(readxl)
library(lme4)
library(ggpp)
library(patchwork)
library(latex2exp)
library(ggdist)
library(ggstance)
library(tidybayes)
library(distributional)
library(posterior)
library(parameters)
library(brms)
library(tidyverse)

# Output options
opts_chunk$set(
  fig.width = 6,
  fig.align = "center",
  eval = TRUE,
  cache = TRUE,
  warning = FALSE,
  error = TRUE,
  message = FALSE
)

# Plotting options
theme_set(
  theme_classic(base_size = 9) +
    theme(
      strip.text = element_text(color = "black", hjust = 0),
      strip.background = element_rect(color = NA, fill = NA),
      strip.text.x = element_text(size = rel(0.9)),
      line = element_line(linewidth = .25),
      plot.tag = element_text(size = rel(1))
    )
)

# Model estimation options
dir.create("models", FALSE)
options(
  mc.cores = 4
)
if ("cmdstanr" %in% installed.packages()[,"Package"]) {
  options(
    brms.backend = "cmdstanr",
    brms.threads = 2
  )
}  
dir.create("data", FALSE)
```

When building and testing theories of psychological phenomena, psychologists have long focused on asking whether an effect exists or not. Yet, establishing that an independent variable affects a dependent variable may not be a sufficient description of the phenomenon if the treatment effect is variable from one individual to another. The importance of such between-person variation, or *heterogeneity*, for theory development is widely recognized, yet rarely communicated sufficiently in the empirical literature [@bolgerCausalProcessesPsychology2019; @brandCausalEffectHeterogeneity2013].

One reason for the scarcity of reporting and interpreting heterogeneity is that psychologists still commonly analyze data with models that obscure the assessment of heterogeneity, such as ANOVA [@bolgerCausalProcessesPsychology2019]. Although more appropriate hierarchical (or multilevel, mixed-effects) models are rapidly becoming more widespread, we believe a second important reason is that researchers do not yet have the conceptual and practical tools required to sufficiently estimate and describe heterogeneity.

When person-to-person variability is modeled and reported, those descriptions often focus on point estimates [@bolgerCausalProcessesPsychology2019], sample statistics [@beyensEffectSocialMedia2020; @vuorreThreeObjectionsNovel2022], or quantities such as the standard deviation of person-specific parameters [@bartosFairCoinsTend2023] which, in isolation, might be difficult to interpret. That is, although we are excited about the increased focus on heterogeneity, we offer that commonly used descriptions of it leave room for improvement.

Because this focus on heterogeneity in psychological phenomena is relatively new, work remains to be done in developing models, descriptions, and reporting practices that maximize investigations' impact on theory development. Our goal is to contribute to that work. We describe and illustrate the use of numerical and graphical descriptions of heterogeneity that (1) Go beyond model parameters to describe heterogeneity in clear and actionable terms, and (2) Take uncertainty in model parameters into account.

Our plan is as follows. We first review established methods for estimating and communicating expected heterogeneity of causal effects in the population using an example dataset from social psychology. We then describe additional ways in which model parameters can be interrogated and transformed to describe distributions of causal effects in the population. We then move beyond summarizing expected degrees of heterogeneity to describing distributions of plausible degrees of heterogeneity. Finally, we extend these methods to compare heterogeneity across different populations using an example dataset from cognitive psychology.

# Review of expected heterogeneity in hierarchical models

To begin our exposition, we reproduce the analyses presented in @bolgerCausalProcessesPsychology2019. In their study, which replicated findings first presented in @scholerInflatingDeflatingSelf2014, 62 participants saw twenty positively and twenty negatively valenced words, and judged whether each word was self-descriptive or not. Because most people are typically motivated to view themselves positively, @bolgerCausalProcessesPsychology2019 predicted that response times to positively valenced words would be shorter than to negatively valenced words [@scholerInflatingDeflatingSelf2014].

```{r data-bolgeretal, include = FALSE}
# Load, save, & clean example data from Bolger et al. (2019)
path <- "data/bolger-etal.zip"
if (!file.exists(path)) {
  download.file(
    "https://github.com/kzee/heterogeneityproject/archive/refs/heads/master.zip",
    destfile = path
  )
  unzip(path, exdir = str_remove(path, ".zip"), junkpaths = TRUE)
}

# Save variables as numerical & factor
dat <- read_csv("data/bolger-etal/heterogeneity_dataset1_traitvalence.csv") |>
  # Cleaning as in Bolger et al 2019
  filter(
    response.keys == "up",
    scale(rt) < 3,
    !(id %in% c(250, 257, 272))
  ) |>
  # Some more cleaning
  mutate(
    person = factor(id),
    trial,
    valence = factor(
      valenceE,
      levels = c(-0.5, 0.5),
      labels = c("Negative", "Positive")
    ),
    rt = logrt
  ) |>
  select(person, trial, valence, rt) |>
  arrange(person, trial)

contrasts(dat$valence) <- c(-0.5, 0.5)
contrasts(dat$valence)
```

## Model 1

```{r}
#| label: tab-dat1
dat |>
  head() |>
  apa_table(
    span_text_columns = FALSE,
    font_size = "small",
    digits = c(0, 0, 0, 2),
    caption = "First six rows of example dataset 1 (Bolger et al., 2019; Exp 1).",
    note = "rt is log-transformed."
  )
```

In this section, we replicate @bolgerCausalProcessesPsychology2019's analysis, using their openly available data (https://github.com/kzee/heterogeneityproject). We first wrangled the data as in @bolgerCausalProcessesPsychology2019, which led to a sample of `r number(nrow(dat), big.mark = ",")` trials from 59 participants that were endorsed as self-relevant. (We present all our code in the online analysis supplement.) We show a sample of these data in Table \@ref(tab:tab-dat1). Then, we estimated the same statistical model (Model 1; equations 1.1-1.3). We modeled the log-transformed reaction time (logRT) of subject $j$ on trial $i$ as a random draw from a normal distribution with mean $\eta$ (*eta*), which could differ between trials $i$ and individuals $j$, and standard deviation $\sigma$ (*sigma*), which we assumed constant across individuals and trials, as indicated by the lack of subscripts:

\begin{equation}
\tag{1.1}
\text{logRT}_{ij} \sim \operatorname{Normal}\left(\eta_{ij}, \sigma^2\right).
\end{equation}

(We note that there are better and more informative alternatives than modeling the log-transformed RTs as normal, but those are outside the scope of this manuscript.) Then, we specified a model of the mean of the logRT distribution ($\eta_{ij}$) such that the regression coefficients captured our substantive questions:

\begin{equation}
\tag{1.2}
\eta_{ij} = \beta_0 + \gamma_{0j} + \left(\beta_1 + \gamma_{1j}\right)\text{V}_{ij}.
\end{equation}

This equation includes two kinds of parameters: $\beta_0$ (*beta*), the intercept, and $\beta_1$, the slope or effect of valence (V), do not have subscripts. In the frequentist tradition, these parameters are considered constants---not modelled on covariates---and typically referred to as "fixed" parameters [e.g. @raudenbushHierarchicalLinearModels2002]. The second set of parameters, $\gamma_{0j}$ (*gamma*) and $\gamma_{1j}$, have the subscript $j$ to indicate that they are person-specific deviations from the average intercept and slope, respectively. That is, $\beta_0 + \gamma_{01}$ is the average reaction time for person $j=1$. In frequentist texts, these are typically called "random" parameters, because they are modeled as varying randomly according to a specified distribution. Following standard multilevel modeling assumptions, we model $\gamma_0$ and $\gamma_1$ as multivariate normal distributed:

\begin{equation}
\tag{1.3}
\begin{bmatrix} 
  \gamma_0 \\ \gamma_1
\end{bmatrix} \sim 
\operatorname{MVN}\left(
  \begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
  \begin{pmatrix} 
    \tau_0 &\rho \\ 
    \rho &\tau_1 
  \end{pmatrix}
\right).
\end{equation}

In this equation, we assume that the subject-specific deviations $\gamma_0$ and $\gamma_1$ have means of zero (because the means are added to them in equation 1.2), standard deviations $\tau$ (*tau*), and a correlation $\rho$ (*rho*). Perhaps confusingly, $\tau$s and $\rho$ are also sometimes called random effects because they describe random (co)variations of the person-specific effects. To be clear, despite this naming convention they are "fixed" features of the population, not of any one group or individual.

What these equations mean substantively is that the extent to which the effect of valence on logRT varies around the average effect ($\beta_1$) is estimated by the standard deviation $\tau_1$. $\tau_0$, on the other hand, describes the standard deviation of the population of individuals' average logRTs across negatively and positively valenced words (intercepts). Moreover, $\rho$ indicates the extent to which individuals' average logRTs correlate with how much their logRTs are affected by valence.

Finally, we contrast coded valence such that negative words were assigned -0.5, and positive words 0.5. This coding results in an intercept that corresponds to the average reaction time across negative and positive words, and a slope that reflects the difference in logRT between negative and positive words.

With data described in Table \@ref(tab:tab-dat1), we can estimate this model using standard maximum likelihood methods as implemented in, for example, the R package lme4 [@batesFittingLinearMixedEffects2015; @rcoreteamLanguageEnvironmentStatistical2023].

```{r fit1-lmer}
fit_lmer <- lmer(
  rt ~ 1 + valence + (1 + valence | person),
  data = dat
)
```

```{r}
# Create a table of coefficients
tab_1 <- model_parameters(fit_lmer, effects = "all") |>
  tibble() |>
  select(Parameter, Coefficient, starts_with("CI_"))

# Bootstrap CIs for variance parameters
path <- "models/lmer-fit1-bootstrap.rds"
if (!file.exists(path)) {
  boots <- confint(
    fit_lmer,
    method = "boot",
    nsim = 1000,
    oldNames = FALSE
  )
  boots <- boots[c(5, 6, 1, 3, 4, 2), ]
  tab_1[, 3:4] <- boots
  saveRDS(tab_1, path)
} else {tab_1 <- readRDS(path)}

# Get approx SEs
tab_1 <- tab_1 |> 
  mutate(SE = (CI_high - Coefficient) / 1.96) |> 
  select(Parameter, Coefficient, SE, CI_low, CI_high)
```

```{r}
#| label: tbl-fit-1
#| cache: false

m1_parnames <- c(
  "$\\beta_0$", "$\\beta_1$",
  "$\\tau_0$", "$\\tau_1$",
  "$\\rho$", "$\\sigma$"
)

tab_1 |>
  mutate(
    mutate(across(where(is.numeric), ~number(.x, .01))),
    CI = str_glue("[{CI_low}, {CI_high}]"),
    Coefficient = ifelse(
      str_starts(Coefficient, "-"), 
      Coefficient, 
      paste0("\\phantom{-}", Coefficient)
    ),
    Parameter = m1_parnames
  ) |> 
  select(
    Parameter, Coefficient, SE, CI
  ) |>
  apa_table(
    span_text_columns = FALSE,
    font_size = "small",
    caption = "Parameter estimates from Model 1.",
    escape = FALSE
  )
mu1n <- tab_1[2, 2, TRUE]
mu1n_low <- tab_1[2, "CI_low", TRUE]
sd1n <- tab_1[4, 2, TRUE]
sd1n_low <- tab_1[4, "CI_low", TRUE]
mu1 <- number(
  as.numeric(tab_1[2, c("Coefficient", "CI_low", "CI_high"), TRUE]), 
  .01
)
sd1 <- number(
  as.numeric(tab_1[4, c("Coefficient", "CI_low", "CI_high"), TRUE]), 
  .01
)
```

We show a traditional summary of the estimated parameters from this model in Table \@ref(tab:tbl-fit-1). For the average person, the estimated effect of valence on logRT is `r mu1[1]` log seconds, with a 95% bootstrap confidence interval (CI) extending from `r mu1[2]` to `r mu1[3]`. The estimated standard deviation of valence effects in the population is `r sd1[1]` log seconds, with a 95% CI of [`r sd1[2]`, `r sd1[3]`].

## Heterogeneity distribution

```{r descriptors, include = FALSE}
# Heterogeneity interval
hi90_num <- qnorm(c(0.05, .95), mu1n, sd1n)
hi90 <- paste(
  number(hi90_num, .01),
  collapse = ", "
)
hi90_low <- paste(
  number(qnorm(c(0.05, .95), mu1n_low, sd1n_low), .01),
  collapse = ", "
)

# Proportions
p_less_zero <- percent(pnorm(0, mu1n, sd1n), .1)
p_rope <- percent(pnorm(0.1, mu1n, sd1n) - pnorm(-0.1, mu1n, sd1n), .1)
# Ratio
ratio <- number(abs(sd1n / mu1n), .01)
ratio_low <- number(abs(sd1n_low / mu1n_low), .01)
```

```{r}
#| label: fig-1
#| fig.height: 1.8
#| fig.env: "figure*"
#| fig.align: "center"
#| fig.cap: Heterogeneity distribution of valence effects and various descriptions of their expected heterogeneity as estimated with Model 1. A. The normal density curve defined by the point estimates of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$). Shaded areas represent areas under the normal curve within 1 (dark) and 2 (light) standard deviations of the mean. B. The 90\% Heterogeneity Interval as represented by a line segment with arrows, and the dark shaded area. C. Proportion of negative valence effects in the population (dark). D. Proportion of valence effects in the population that are within the region of practical equivalence to zero (ROPE; dark).

fig_0 <- ggplot() +
  aes(xdist = dist_normal(mu1n, sd1n)) +
  scale_x_continuous(breaks = extended_breaks()) +
  coord_cartesian(xlim = c(-0.55, 0.25)) +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_text(lineheight = rel(0.5))
  ) +
  scale_fill_manual(values = c("grey90", "grey80"), na.value = "white")
d1 <- dist_normal(mu1n, sd1n) |>
  mean_qi(.width = c(.66, .9))

fig_1 <- fig_0 +
  stat_slab(
    aes(fill = after_stat(level)),
    .width = c(.68, .95),
    p_limits = c(.000001, .999999),
    slab_color = "grey10",
    slab_linewidth = .4
  ) +
  geom_text(
    label = TeX("$\\tau_1$"),
    x = mu1n + .05, y = .175,
    size = 3,
    parse = TRUE
  ) +
  annotate(
    "segment",
    x = mu1n, xend = mu1n + sd1n, y = .1, yend = .1,
    linewidth = .35,
    arrow = arrow(length = unit(4, "pt"), type = "closed")
  ) +
  geom_text(
    label = TeX("$\\beta_1$"),
    x = mu1n - .075, y = .075,
    size = 3,
    parse = TRUE
  ) +
  annotate(
    "segment",
    x = mu1n, xend = mu1n, y = .1, yend = 0,
    linewidth = .35,
    arrow = arrow(length = unit(4, "pt"), type = "closed")
  ) +
  theme(axis.title.y = element_text(angle = 90))

fig_2 <- fig_0 +
  stat_slab(
    aes(fill = after_stat(level)),
    .width = c(.90, 1),
    p_limits = c(.000001, .999999),
    slab_color = "grey10",
    slab_linewidth = .4
  ) +
  geom_text(
    label = TeX("$HI_{90}$"),
    x = mu1n, y = .175,
    size = 3,
    parse = TRUE
  ) +
  annotate(
    "segment",
    x = d1$.lower[2], xend = d1$.upper[2], y = .1, yend = .1,
    linewidth = .35,
    arrow = arrow(length = unit(4, "pt"), type = "closed", ends = "both")
  )

fig_3 <- fig_0 +
  stat_slab(
    aes(fill = after_stat(x < 0)),
    p_limits = c(.000001, .999999),
    slab_color = "grey10",
    slab_linewidth = .4
  ) +
  geom_text(
    label = TeX("$p^-$"),
    x = mu1n, y = .175,
    size = 3,
    parse = TRUE
  )

fig_4 <- fig_0 +
  stat_slab(
    aes(fill = after_stat(between(x, -.1, .1))),
    p_limits = c(.000001, .999999),
    slab_color = "grey10",
    slab_linewidth = .4
  ) +
  geom_text(
    label = TeX("$p^{ROPE}$"),
    x = 0.0, y = 0.5, hjust = 0,
    size = 3,
    parse = TRUE
  )

(fig_1 | fig_2 | fig_3 | fig_4) &
  scale_y_continuous(
    expand = expansion(c(0.0, 0.05))
  ) &
  labs(
    y = "Density",
    x = "Valence effect"
  ) &
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) &
  plot_annotation(tag_levels = "A")
```

Rows 2 and 4 in Table \@ref(tab:tbl-fit-1) define the expected normal distribution of valence effects in the population, visualized in Figure \@ref(fig:fig-1)A. In other words, our point estimate of the distribution of valence effects is Normal(`r mu1[1]`, `r sd1[1]`^2^). However, it is an incomplete description of heterogeneity for two reasons. First, it does not necessarily communicate the degree of heterogeneity in clear and actionable terms: In addition to the SD, we want quantities that describe heterogeneity in clearer terms. Below, we introduce several metrics that e.g. directly describe where a given proportion of the slopes are expected to fall. Second, it ignores the uncertainty with which the heterogeneity parameters are estimated. That is, we need ways to describe (uncertainty) distributions of distributions.

We now turn to this first issue and outline several additional metrics of heterogeneity afforded by the model's parameter estimates.

### Interval descriptors

First, we can use the point estimates in Table \@ref(tab:tbl-fit-1) to construct an expected *heterogeneity interval* that describes the range within which a certain percentage of the population's slopes are expected to fall [@bolgerCausalProcessesPsychology2019]. To do so, we must first determine an appropriate percentage to describe: By convention, @bolgerCausalProcessesPsychology2019 and others have focused on the 95% heterogeneity interval ($HI_{95}$). However, because there are already confusingly many quantities using the five percent cutoff, in this manuscript we focus on the 90% heterogeneity interval, and reserve 95% to describing uncertainties. The appropriate percentage to describe with a heterogeneity interval is determined by the substantive and communicative aims at hand; for our illustration 90% seemed reasonable.

To calculate a heterogeneity interval, we first specify the desired probability limits. For a 90% interval, we use .05 and .95, which together define the central 90% of the distribution. Then, we pass those probabilities and the estimated mean and standard deviation to the normal quantile function $\Phi^{-1}$ (*phi*, `qnorm()` in R), to obtain the interval: $HI_{90} = \Phi^{-1}([.05, .95], \beta_1, \tau_1) = \Phi^{-1}([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`]. In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval. We illustrate this interval in Figure \@ref(fig:fig-1)B.

### Proportion descriptors

In contrast, proportions of slopes above or below some critical value, or within some critical range, might serve as more informative summaries. For example, we might ask "What proportion of individuals in the population respond faster to positively valenced words?" In other words, we ask what proportion of the heterogeneity distribution is below zero. To answer, we pass zero and the estimated mean and standard deviation to the normal cumulative distribution function ($\Phi$; `pnorm()` in R): $p^- = \Phi(0, \beta_1, \tau_1) = \Phi(0,$ `r mu1[1]`, `r sd1[1]`) = `r p_less_zero`. This number---$p^-$ for proportion of population with negative effects---is the probability that a random slope from this population would take a negative value, or, in other words, the proportion of individuals in the population with negative valence effects. We illustrate this probability in Figure \@ref(fig:fig-1)C.

However, using zero as a critical value might not be sufficiently informative, especially when theory allows specifying a smallest effect size of interest, or what is known as a region of practical equivalence [ROPE, @anvariUsingAnchorbasedMethods2021; @lakensEquivalenceTestingPsychological2018; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014]. In common applications, ROPE is used to statistically infer whether an estimated parameter, such as the effect of valence on logRT for the average person, is practically significant. But we can equally well use a theory-informed region of parameter values to describe and make inferences about the heterogeneity distribution of this effect in the population.

For example, we might know from theory that valence effects in the interval [-0.1, 0.1] are practically equivalent to zero. To calculate, we can again use the normal cumulative distribution function to calculate the proportion of individuals in the population whose valence effect falls within this interval or region of practical equivalence: $p^{ROPE} = \Phi(0.1, \beta_1, \tau_1) - \Phi(-0.1, \beta_1, \tau_1)$ = `r p_rope`. In words, `r p_rope` of the population is expected to have valence effects that are practically equivalent to zero. Note that this statement's validity critically depends on the chosen interval, which we arbitrarily set as [-0.1, 0.1] in this example. We visualize this probability in Figure \@ref(fig:fig-1)D.

### Ratio descriptors

Although the interval and proportion descriptors show where the population's slopes are likely to fall, they do so in absolute terms such as log(rt) in the running example. A contrasting or *relative* way to describe heterogeneity is to express it as a ratio of the standard deviation to the mean. The benefit of this descriptor is that in isolation a standard deviation says very little, but when directly compared to the mean effect, it allows for a sufficient description of the relative degree of heterogeneity of the phenomenon under study.

[@bolgerCausalProcessesPsychology2019, p.609] suggest as a rule of thumb that if the ratio of the standard deviation to the average effect is 0.25 or greater, then heterogeneity can be deemed noteworthy. With these data and model, the ratio $\frac{\tau_1}{\beta_1}$ is `r ratio`, suggesting that the degree of heterogeneity in valence effects is noteworthy. (We reiterate the point made by @bolgerCausalProcessesPsychology2019 that the cutoff value of 0.25 is arbitrary and researchers should choose the cutoff based on their substantive goals.)

## Uncertainty

Although informative, the expected normal distribution of valence effects and its transformations, such as the ones illustrated in Figure \@ref(fig:fig-1), ignore the uncertainty inherent in the estimated parameters. The point estimates alone cannot fully communicate the model's knowledge about the degree of heterogeneity. That is, the uncertainty with which we have estimated the population parameters, indicated by e.g. the confidence intervals in Table \@ref(tab:tbl-fit-1), is not carried forward to the calculations described above or any description based on them.

For example, we only presented the expected heterogeneity interval, but not the range of credible values that it might take. If we re-calculate $HI_{90}$ based on the lowest 95%CI bounds in Table \@ref(tab:tbl-fit-1), the resulting heterogeneity interval indicates that 90% of the population's slopes would fall in the [`r hi90_low`] interval. In contrast to the expected interval, this interval implies that 90% of the population's slopes are expected to fall below zero.

We have now arrived at the crux of the current work: How should we estimate and describe heterogeneity in psychological phenomena such that the fundamental uncertainty in the estimated parameters is retained?

# Probabilistic assessment of heterogeneity

We think that probabilistic (i.e. "bayesian") methods are uniquely able to address this challenge. Bayesian methods estimate and describe heterogeneity in ways that naturally incorporate information about uncertainty because they deliver distributions of plausible values that could underlie hypothetical data generating processes [@gelmanBayesianDataAnalysis2013; @kruschkeBayesianDataAnalysis2017; @kruschkeDoingBayesianData2014; @mcelreathStatisticalRethinkingBayesian2020]. For typical scenarios, probabilistic models are as easy to use as their maximum likelihood counterparts [@burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018] Moreover, many advanced models *require* bayesian methods [@muthenBayesianStructuralEquation2012; @asparouhovDynamicStructuralEquation2018].

```{r fit1-brms}
fit_1 <- brm(
  rt ~ 1 + valence + (1 + valence | person),
  data = dat,
  file = "models/brm-fit-1",
  control = list(adapt_delta = .95)
)
```

```{r}
#| label: ppcheck-1
#| eval: false

# Quick model checking
pairs(fit_1$fit, pars = variables(fit_1)[1:6])
mcmc_plot(fit_1, type = "rhat_hist")
pp_check(fit_1)
```

```{r}
#| label: tbl-fit-1-brms
#| cache: false

tab1 <- parameters(fit_1, centrality = "mean", effects = "all") |>
  tibble() |>
  select(Parameter, Mean, starts_with("CI_"))
```

```{r}
#| label: tbl-samples-1
#| cache: false

post1 <- as.data.frame(fit_1, variable = c("b_", "sd_", "cor_", "sigma"), regex = TRUE) |>
  tibble() |>
  rownames_to_column("Sample")
post1_sum <- fit_1 |>
  parameters(effects = "all", centrality = "mean", dispersion = TRUE) |>
  tibble() |>
  mutate(
    Mean = number(Mean, .01),
    SD = number(SD, .001)
  ) |>
  select(Parameter, Mean, SD) |>
  pivot_longer(c(Mean, SD), names_to = "Sample") |>
  pivot_wider(names_from = Parameter, values_from = value)

set.seed(191)
post1 |>
  slice_sample(n = 6) |>
  arrange(as.integer(Sample)) |>
  mutate(
    across(-Sample, ~ number(., .01))
  ) |>
  add_row(post1_sum) |>
  setNames(c("Sample", m1_parnames)) |>
  apa_table(
    span_text_columns = FALSE,
    font_size = "small",
    midrules = 6,
    caption = "Six random samples from Model 1's parameters' posterior distributions, and their summaries.",
    escape = FALSE,
    placement = "h"
  )
```

The goal of bayesian inference is the multivariate posterior probability distribution of the model's parameters. However, because closed-form solutions are not available for posterior distributions of many important types of statistical models, in practice bayesian methods rely on drawing many random samples from the posterior distribution [@gelmanBayesianDataAnalysis2013; @ravenzwaaijSimpleIntroductionMarkov2016]. Therefore, "estimate" means a large number of random samples from the posterior distribution. These samples can then be used to calculate, summarize, and visualize the posterior distribution of any desired quantity.

In practice, one obtains (for example) 4,000 samples from the posterior distribution using computational algorithms [e.g. @standevelopmentteamStanModelingLanguage2023] through accessible software [e.g., @burknerBrmsPackageBayesian2017], and then summarizes them using familiar data processing techniques [e.g. @rcoreteamLanguageEnvironmentStatistical2023]. Here, we used the R package brms [@burknerBrmsPackageBayesian2017] to specify the model and then draw random samples from its posterior distribution. The MCMC estimation algorithm completed in about 5 seconds on a modern laptop. We then assessed the estimation algorithm convergence graphically and numerically, and model adequacy using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013].

Table \@ref(tab:tbl-samples-1) shows six random samples from the posterior distribution of Model 1's population-level parameters ("fixed" parameters in the frequentist nomenclature). The two bottom rows show their means and standard deviations (which correspond to frequentist standard errors). Note that because we used brms's default noninformative prior distributions, the posterior summaries are numerically very similar to the maximum likelihood estimates in Table \@ref(tab:tbl-fit-1).

## Heterogeneity distribution

Armed with the bayesian estimates (Table \@ref(tab:tbl-samples-1)), we now return to the distribution of valence effects in the population. We now have 4,000 samples from this heterogeneity distribution's posterior distribution. We first redraw the expected heterogeneity distribution from Figure \@ref(fig:fig-1) using the posterior mean values of $\beta_1$ and $\tau_1$ in Figure \@ref(fig:fig-2)A (thick dark blue curve). Superimposed on that normal density curve are heterogeneity distributions calculated not from the posterior means, but from 100 random posterior samples of $\beta_1$ and $\tau_1$. From these curves we glean that the distribution of valence effects might well be narrower (i.e. less heterogeneity) or wider (more heterogeneity) than is suggested by the point estimates.

```{r}
#| label: calc-dfs

set.seed(99)
x <- seq(-0.6, 0.3, length = 301)

calculate_density <- function(posterior, b, sd) {
  posterior |>
    crossing(x) |>
    mutate(
      pdf = dnorm(x, {{b}}, {{sd}}),
      cdf = pnorm(x, {{b}}, {{sd}})
    )
}

post1_density_mean <- post1 |>
  mean_qi(b_valence1, sd_person__valence1) |>
  calculate_density(b_valence1, sd_person__valence1)

post1_density_samples <- post1 |>
  slice_sample(n = 100) |>
  calculate_density(b_valence1, sd_person__valence1)

post1_density_ribbon <- post1 |>
  select(b_valence1, sd_person__valence1) |>
  calculate_density(b_valence1, sd_person__valence1) |> 
  group_by(x) |>
  curve_interval(.width = 0.9)
```

```{r}
#| label: person-specific-valence-effects
#| eval: false

# Just sketching a figure of person-specific valence slopes
coef(fit_lmer)$person |> 
  as.data.frame() |> 
  rownames_to_column("person") |> 
  left_join(
    coef(fit_1)$person |> 
      as.data.frame() |> 
      rownames_to_column("person"),
    by = join_by(person)
) |> 
  select(person, MLE = valence1, Bayes = Estimate.valence1) |> 
  pivot_longer(-person) |> 
  ggplot(aes(value, name)) +
  geom_line(aes(group = person)) +
  geom_point()
summary(fit_1)
```

```{r}
#| label: fig-2
#| fig.height: 3
#| fig.width: 5.4
#| fig.env: "figure*"
#| fig.cap: "Bayesian estimates of the heterogeneity distribution of valence effects. A. Probability density function (PDF) curves. The thick line illustrates the expected (point estimate; posterior mean) PDF of the valence effect heterogeneity distribution. Thin lines show 100 random samples of the PDF's posterior distribution, and facilitate visual communication of uncertainty in the PDF and, therefore, the valence effects' heterogeneity distribution. Dark line segments are inside the [-0.1, 0.1] ROPE, light segments are outside. The y-axis is in arbitrary probability density units at each point of the x-axis. B. The heterogeneity distribution's posterior mean PDF curve (line), and 90\\%CI (ribbon). C. Cumulative density function (CDF) curves, annotated as A. D. Histogram of the posterior distribution of $p^-$ (CDF segments below zero): This value describes the proportion of individuals with negative valence effects on logRT. Point and interval representes the posterior mean and 95\\%CI. E. As C, but represents uncertainty with a 95\\% confidence ribbon. F. Approximate posterior density of $p^-$."

# Shared Y axis
scale_y_01 <- scale_y_continuous(
  breaks = extended_breaks(7),
  expand = expansion(c(0.005, 0.005)),
  limits = c(0, 1)
)

# Posterior samples of pdf and cdf curves
p_cdf_samples <- post1_density_mean |>
  ggplot() +
  aes(x, cdf) +
  ylab("Cumulative density") +
  scale_color_manual(
    values = c("dodgerblue1", "dodgerblue4")
  ) +
  scale_x_continuous(
    "Valence effect",
    expand = expansion(0),
    breaks = extended_breaks()
  ) +
  scale_y_01 +
  geom_vline(
    xintercept = 0, 
    lty = "dotted", 
    linewidth = .25
  ) +
  geom_line(
    linewidth = .6,
    col = "dodgerblue4"
  ) +
  theme(legend.position = "none")

p_pdf_samples <- p_cdf_samples +
  aes(y = pdf) +
  geom_line(
    data = post1_density_samples,
    aes(group = Sample),
    col = "dodgerblue1",
    alpha = .33,
    linewidth = .15
  ) +
  scale_y_continuous(
    breaks = extended_breaks(7),
    expand = expansion(c(0.005, 0.005)),
  ) +
  ylab("Probability density") +
  theme(
    axis.title.x = element_blank()
  )

p_cdf_samples <- p_cdf_samples +
  geom_line(
    data = post1_density_samples,
    aes(group = Sample, color = after_stat(between(x, -0.1, 0.1))),
    alpha = .33,
    linewidth = .15
  )

# Posterior mean and 95%CI pdf and cdf curves
p_cdf_ribbon <- post1_density_ribbon |>
  ggplot(aes(x = x)) +
  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper) +
  scale_x_continuous(
    "Valence effect",
    expand = expansion(0),
    breaks = extended_breaks()
  ) +
  scale_y_continuous(
    breaks = extended_breaks(7),
    expand = expansion(c(0.005, 0.005))
  ) +
  geom_vline(
    xintercept = 0, 
    lty = "dotted", 
    linewidth = .25
  ) +
  geom_lineribbon(
    linewidth = .5,
    color = "dodgerblue4",
    fill = alpha("dodgerblue1", .33)
  ) +
  theme(
    legend.position = "none",
    axis.title.y = element_blank()
  )

p_pdf_ribbon <- p_cdf_ribbon +
  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
  theme(
    axis.title.x = element_blank(), 
  )

# Add summaries of CDF < 0 to CDF curves as inset plots
p_p0 <- post1 |>
  mutate(dist = dist_normal(b_valence1, sd_person__valence1)) |>
  ggplot() +
  scale_x_continuous(
    expand = expansion(c(0.0, 0.1))
  ) +
  scale_y_01 +
  theme_void(base_size = 9) +
  theme(
    plot.tag.position = c(.55, .75), 
    plot.tag = element_text(size = rel(1))
  )
p_p0_samples <- p_p0 + 
  stat_histinterval(
    aes(y = cdf(dist, 0)),
    breaks = 50,
    fill = alpha("dodgerblue1", 0.5),
    color = "dodgerblue4",
    side = "right",
    .width = .95,
    size = 1,
    scale = 0.5,
    justification = 0.025
  )
p_p0_density <- p_p0 + 
  stat_slabinterval(
    aes(y = cdf(dist, 0)),
    fill = alpha("dodgerblue1", 0.5),
    color = "dodgerblue4",
    side = "right",
    .width = .95,
    size = 1,
    scale = 0.5,
    justification = 0.025
  )
p_cdf_samples <- p_cdf_samples +
  inset_element(
    p_p0_samples,
    0, 0, .2, 1, 
    align_to = "panel", 
    on_top = TRUE
  )
p_cdf_ribbon <- p_cdf_ribbon +
  inset_element(
    p_p0_density,
    0, 0, .2, 1, 
    align_to = "panel", 
    on_top = TRUE
  )

p_cdf <- p_cdf_samples | p_cdf_ribbon

p_pdf <- p_pdf_samples | p_pdf_ribbon

(p_pdf / p_cdf) +
  plot_annotation(tag_levels = "A")
```

Moreover, the location of these curves differ: Some curves are further to the left (valence effect for the average person is more negative), and some further to the right (effect for the average person is more positive). The distribution of these curves represents our current knowledge about the heterogeneity distribution of valence effects in the population---given these data and this model---and can be more completely summarized graphically as a ribbon: Figure \@ref(fig:fig-2)B shows the 95%CI of the valence effect's probability density for each value of the effect on the x-axis. A sufficient description of heterogeneity must include information about uncertainty in both the location (mean) and scale (standard deviation) parameters of the heterogeneity distribution.

However, depicting the heterogeneity distribution visually as a probability density function (PDF) curve has its drawbacks. First, it appears to us visually more challenging to read the degree of uncertainty from a PDF, even when the underlying parameters' uncertainty is summarized as a ribbon around the posterior mean (Figure \@ref(fig:fig-2)B). Second, for many applications, the y-axis is not informative: We do not care that the probability density of the curve is (for example) 3.0 at some specific value of the valence effect.

Therefore, in Figure \@ref(fig:fig-2)C we depict the heterogeneity distribution as a *cumulative density* function (CDF). We again superimposed 100 random samples from the curve's posterior distribution on the posterior mean, and summarize the entire posterior distribution with a 95% credibility ribbon in Figure \@ref(fig:fig-2)D. We believe the CDF is particularly useful as a descriptor of heterogeneity, because the y-axis describes a directly interpretable quantity: Namely, the proportion of population with valence effects below some specific value.

### Interval descriptors

```{r}
p1n <- post1 |>
  mutate(
    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
  ) |>
  mean_qi(hi_low, hi_high)
p1s <- p1n |>
  mutate(
    across(where(is.numeric), ~ number(., .01)),
    l = str_glue("[{hi_low.lower}, {hi_low.upper}]"),
    u = str_glue("[{hi_high.lower}, {hi_high.upper}]")
  )
```

Above, we described the heterogeneity interval as a range of values where a specific percentage of the population's slopes are expected to fall (e.g. $HI_{90}$ for a 90% heterogeneity interval). However, a single interval cannot accommodate the uncertainty with which the underlying parameters are estimated. To carry uncertainty forward from model parameters to their transformations, such as the $HI_{90}$, we redo the calculations from above, but instead of using only the mean's and standard deviation's point estimates, we repeat the calculations for each of the 4,000 randomly sampled pairs of $\beta_1$ and $\tau_1$. Consequently, we get 4,000 samples from the $HI_{90}$s posterior distribution (Figure \@ref(fig:intervals)).

Summarizing a distribution of intervals entails some challenges, however, because an interval is defined by two quantities---the lower and upper bounds. The 95% most credible lower bounds of $HI_{90}$ range between `r p1s$l`, whereas the 95% most credible upper bounds range between `r p1s$u` (Figure \@ref(fig:intervals)B). Thus, to adequately describe an estimated heterogeneity interval, researchers must communicate two separate uncertainty intervals: In words, we estimate that 90% of the population's valence effects range from `r paste(p1s$hi_low, p1s$l)` to `r paste(p1s$hi_high, p1s$u)`.

Figure \@ref(fig:intervals)A further suggests that communicating the two uncertainty intervals of a heterogeneity interval is not only cumbersome, but also ignores potential correlations between the heterogeneity interval endpoints' posterior distributions. For these reasons, although the HI can be a useful summary, we occasionally favor [e.g., @vuorreAffectiveUpliftVideo2023] the scalar descriptors discussed below.

### Proportion descriptors

A complementary description of heterogeneity is the proportion of the population whose effects fall above or below some critical value. For example, we can calculate proportions with negative and positive effects by using zero as the critical value. In this example, we ask "What proportion of individuals in the population have negative effects of valence?"

```{r}
tmp <- post1 |>
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1),
    pd = cdf(dist, 0),
    rope = cdf(dist, 0.1) - cdf(dist, -0.1),
    ratio = abs(sd_person__valence1 / b_valence1)
  ) |>
  mean_qi(pd, rope, ratio) |>
  mutate(
    pd_r = str_glue(
      "{percent(pd, .1)} [{percent(pd.lower, .1)}, {percent(pd.upper, .1)}]"
    ),
    pd_r_n = str_glue(
      "{percent(1-pd, .1)} [{percent(1-pd.upper, .1)}, {percent(1-pd.lower, .1)}]"
    ),
    rope_r = str_glue(
      "{percent(rope, .1)} [{percent(rope.lower, .1)}, {percent(rope.upper, .1)}]"
    ),
    rope_r_n = str_glue(
      "{percent(1-rope, .1)} [{percent(1-rope.upper, .1)}, {percent(1-rope.lower, .1)}]"
    ),
    ratio_r = str_glue(
      "{number(ratio, .01)} [{number(ratio.lower, .01)}, {number(ratio.upper, .01)}]"
    )
  )
```

To answer, we calculate $p^- = \Phi(0, \beta_1, \tau_1)$ for each posterior sample of $\beta_1$ and $\tau_1$. We show 100 posterior samples of the CDF in Figure \@ref(fig:fig-2)C with a vertical line superimposed at zero. The y-axis value where the CDF crosses zero on the x-axis indicates the population proportion of negative valence effects ($p^-$). We also show a histogram of all 4,000 posterior samples of that proportion in the top left margin of Figure \@ref(fig:fig-2)C, with the associated 95%CI. The model predicts the proportion of individuals in the population with negative valence effects to be approximately `r percent(as.numeric(tmp$pd), .1)` (posterior mean), but with 95% confidence this value could be as low as `r percent(as.numeric(tmp$pd.lower), .1)` or as high as `r percent(as.numeric(tmp$pd.upper), .1)`. Alternatively, the model predicts that `r tmp$pd_r_n` of individuals in the population would show reversals of the valence effect.

Moreover, if theory allows defining a range of parameter values that are practically equivalent to zero (ROPE), we can use the posterior distribution to quantify uncertainty in the proportion of individuals predicted to have such practically negligible effects. We colored the curves in Figure \@ref(fig:fig-2)C such that curve segments within [-0.1, 0.1] are highlighted with a darker color. Those darker line segments represent proportions of the population whose valence effect is practically equivalent to zero ($p^{ROPE}$). To quantify uncertainty in $p^{ROPE}$ we then aggregate the segments' to a mean and a 95%CI: `r tmp$rope_r` of individuals in the population have a valence effect that is practically equivalent to zero. We note that the ROPE of [-0.1, 0.1] here was arbitrary and picked just to illustrate the example.

So far, these examples have highlighted the importance of quantifying uncertainty in descriptions of heterogeneity. Had we only focused on the point estimates (posterior means), we might have misleadingly concluded that $p^-$ = `r percent(tmp$pd, .1)` and $p^{ROPE}$ = `r percent(tmp$rope, .1)`. However, with 95% confidence, these values might be as small as `r percent(tmp$pd.lower, .1)` and `r percent(tmp$rope.lower, .1)`, or as large as `r percent(tmp$pd.upper, .1)` and `r percent(tmp$rope.upper, .1)`, respectively.

### Ratio descriptors

Finally, we can assess heterogeneity in relative terms by comparing the magnitude of the heterogeneity in valence effects (the standard deviation $\tau_1$) to the magnitude of the average effect (the mean $\beta_1$) by calculating the ratio $\frac{\tau_1}{\beta_1}$. Figure \@ref(fig:hdist-5)A shows 4,000 samples from the joint posterior distribution of the mean and standard deviation, from which we calculated 4,000 samples of the posterior distribution of $\frac{\tau_1}{\beta_1}$ (Figure \@ref(fig:hdist-5)B). The ratio `r tmp$ratio_r` suggests that the relative magnitude of heterogeneity is substantial, but might be as low as `r number(tmp$ratio.lower, .01)` or as great as `r number(tmp$ratio.upper, .01)`, with 95% confidence. If we used the 1/4 rule of thumb suggested in @bolgerCausalProcessesPsychology2019, with these results we could say *with confidence* that heterogeneity in valence effects is notable.

```{r}
#| label: hdist-5
#| fig.height: 2.2
#| fig.width: 4
#| fig.cap: Panel A. 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$). B. Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\%CI.

p1 <- post1 |>
  ggplot(aes(b_valence1, sd_person__valence1)) +
  scale_y_continuous(
    TeX("$\\tau_1$")
  ) +
  scale_x_continuous(
    TeX("$\\beta_1$")
  ) +
  geom_point(
    shape = 1,
    col = "dodgerblue1",
    alpha = .25,
    size = .75
  )
p2 <- post1 |>
  mutate(ratio = abs(sd_person__valence1 / b_valence1)) |>
  ggplot(aes(ratio)) +
  scale_y_continuous(
    "Count",
    expand = expansion(0)
  ) +
  scale_x_continuous(
    TeX("$\\frac{\\tau_1}{\\beta_1}$"),
    breaks = extended_breaks()
  ) +
  stat_histinterval(
    breaks = 100,
    fill = alpha("dodgerblue1", 0.5),
    color = "dodgerblue4",
    .width = .95,
    size = 1,
    justification = 0.01
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
  )
(p1 | p2) &
  theme(
    aspect.ratio = 1
  ) &
  plot_annotation(tag_levels = "A")
```

We have seen that---with these example data and this model---our uncertainty in the estimated heterogeneity metrics is substantial: Point estimates provide at best incomplete descriptions of our current state of knowledge regarding heterogeneity in valence effects. We will next see that incorporating uncertainty is not only useful but critical when we turn from describing heterogeneity in one population to comparing its magnitude across multiple populations.

# Comparing heterogeneity

Next, we move beyond assessing heterogeneity in one population to comparing degrees of heterogeneity across potentially multiple populations of different kinds of study units. To illustrate, we reanalyze a dataset from @mahVariabilitySubjectsFree2023 addressing differences in between person variability (heterogeneity) in memory performance between a free recall memory experiment and a cued recall memory experiment. In @mahVariabilitySubjectsFree2023's Experiment 3, 260 individuals studied a list of twenty target words. After a short break, they then either freely recalled as many of the target words as they could (Free recall group, N = 123) or recalled as many target words as they could when prompted with related cue words (Cued recall group, N = 137). Thus, the Free and Cued recall tasks had different groups of participants but the same target words. We show a sample of these data in Table \@ref(tab:tab-dat2).

```{r data-mah-lindsay-3, include = FALSE}
# Experiment 3: task type is between-person
path <- "data/mah-lindsay-experiment-3.zip"
if (!file.exists(path)) {
  download.file(
    "https://files.ca-1.osf.io/v1/resources/pfhu9/providers/osfstorage/?zip=",
    destfile = path
  )
  unzip(path, exdir = str_remove(path, ".zip"), junkpaths = TRUE)
}

dat <- read_excel(
  "data/mah-lindsay-experiment-3/CuedVsFree_DRM_2022_Item_Data.xlsx"
) |>
  filter(!Exclude) |>
  mutate(
    person = factor(idnumber),
    task = factor(test_type, levels = c("FR", "CR"), labels = c("Free", "Cued")),
    accuracy
  ) |>
  select(person, task, target, accuracy) |>
  # Two observations were miscoded
  mutate(accuracy = if_else(accuracy == 2, 1, accuracy)) |> 
  arrange(task, person, target)
```

With a preregistered Pitman-Morgan test, @mahVariabilitySubjectsFree2023 found that participants who completed the cued recall task showed greater heterogeneity in memory performance than did the free recall group: The Cued:Free recall variance ratio was 1.33 (with a [1.14, 1.54] 95% bootstrap interval). @mahVariabilitySubjectsFree2023 also confirmed this result by comparing multilevel models that did and didn't allow for different between-person variabilities in memory ability for the two groups, and found similar results across three different experiments.

```{r tab-dat2}
dat |>
  slice(4:6, .by = task) |> 
  apa_table(
    span_text_columns = FALSE,
    font_size = "small",
    escape = FALSE,
    digits = 0,
    caption = "Six rows of example dataset 2 (Mah \\& Lindsay, 2023; Exp 3)."
  )
```

We extend those analyses to illustrate how our descriptions of heterogeneity generalize to differences in heterogeneity in ways that inform audiences beyond variance ratios and model comparisons. We ask three questions about between-task differences in heterogeneity: (1) To what extent is memory performance more variable between people in the cued recall task compared to the free recall task? (2) To what extent is memory performance variability between target words different across cued recall and free recall tasks? And (3) How consistent is target word heterogeneity across the two tasks: Are target words associated with good memory performance in cued recall experiments the same words that are associated with good memory performance in free recall experiments?

To answer these questions, we model the $i$th total recall accuracy in 1 to `r nrow(dat)`, of person $j$ in 1 to `r length(unique(dat$person))`, word $k$ in 1 to `r length(unique(dat$target))`, and task $m$ in {F (free recall), C (cued recall)} as Bernoulli distributed, where the probability of an accurate answer is determined by the rate parameter $\pi_{ijkm}$. As is common with generalized linear models, we model the rate parameter through a nonlinear link function. In this example, we use the cumulative normal density function ($\Phi$, or "probit"), but other link functions are also available, such as the common logit. Consequently, it is the "linear predictor" $\eta$ that we then model as a linear combination of the predictors.

\begin{align*}
\tag{2}
\text{Accuracy}_{ijkm} &\sim \operatorname{Bernoulli}\left(\pi_{ijkm}\right) \\
\pi_{ijkm} &= \Phi\left(\eta_{ijkm}\right) \\
\eta_{ijkm} &= \beta_{m} + \gamma_{jm} + \delta_{km} \\
\gamma_{[m:F]} &\sim 
  \operatorname{Normal}\left(0, \tau_{\gamma_{[m:F]}} \right) \\
\gamma_{[m:C]} &\sim 
  \operatorname{Normal}\left(0, \tau_{\gamma_{[m:C]}} \right) \\
\begin{bmatrix} 
  \delta_{[m:F]} \\ \delta_{[m:C]}
\end{bmatrix} &\sim 
  \operatorname{MVN}\left(
  \begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
  \begin{pmatrix} 
    \tau_{\delta_{[m:F]}} &\rho_\delta \\ 
    \rho_\delta &\tau_{\delta_{[m:C]}} 
  \end{pmatrix}
\right).
\end{align*}

This Model 2 of memory performance is similar to our Model 1 of valence effects above, but contains two sources of heterogeneity (persons, whose parameters we represent with $\gamma$, and target words, whose parameters we write with $\delta$). In addition, instead of coding the task type (free recall vs. cued recall) using predictor coding schemes such as contrast or dummy coding, we have index-coded task using subscripts $_{m:F}$ to stand for Free recall parameters, and $_{m:C}$ for parameters pertaining to the Cued recall task. This reparameterization is useful, because it allows us to quantify heterogeneity in memory performance separately for the two tasks, rather than for (if using contrast coding) the average task and their difference. Notice also that we model $\gamma$ using to independent normal distributions, and $\delta$ with a multivariate normal distribution. Because different persons participated in the two tasks, we cannot assess whether participant-specific abilities are correlated across the tasks. But we can assess this for target items, which were common across the tasks.

```{r fit-brms-ml3}
#| message: false

model <- bf(
  accuracy ~ 0 + task + 
    (1 | gr(person, by = task)) + 
    (0 + task | target) 
)
fit2 <- brm(
  model,
  family = bernoulli("probit"),
  data = dat,
  silent = 0,
  file = "models/brm-fit-ml3-target",
  control = list(adapt_delta = .95)
)
```

```{r}
#| label: ppcheck-2
#| eval: false

# Quick model checking
pairs(fit2$fit, pars = variables(fit2)[1:7])
mcmc_plot(fit2, type = "rhat_hist")
pp_check(fit2, type = "bars_grouped", group = "task", ndraws = 100)
```

We estimated Model 2 exactly as Model 1, by drawing 4,000 random samples from its posterior distribution [@burknerBrmsPackageBayesian2017]. We then confirmed graphically and numerically that the estimation algorithm had converged, and that the model performed adequately using a graphical posterior predictive check [@gelmanBayesianDataAnalysis2013]. We summarise the model's posterior distribution in Table \@ref(tab:tbl-samples-2).

```{r}
#| label: tbl-samples-2
#| cache: true

m2_parnames <- c(
  "$\\beta_{m:F}$", "$\\beta_{m:C}$",
  "$\\tau_{\\gamma_{m:F}}$", "$\\tau_{\\gamma_{m:C}}$",
  "$\\tau_{\\delta_{m:F}}$", "$\\tau_{\\delta_{m:C}}$",
  "$\\rho_\\delta$"
)

post2 <- as_draws_df(
  fit2, 
  variable = c("b_", "sd_", "cor_"), 
  regex = TRUE
) |> 
  select(-.iteration, -.chain) |> 
  relocate(.draw, .before = b_taskFree)

set.seed(191)
post2 |>
  slice_sample(n = 6) |>
  mutate(.draw = as.character(.draw)) |> 
  add_row(
    post2 |> 
      summarise_draws(
        Mean = ~mean(.),
        SD = ~sd(.)
      ) |> 
      pivot_longer(c(Mean, SD), names_to = ".draw") |>
      pivot_wider(names_from = variable, values_from = value)
  ) |> 
  setNames(c("Sample", m2_parnames)) |> 
  apa_table(
    span_text_columns = FALSE,
    font_size = "scriptsize",
    midrules = 6,
    digits = 2,
    caption = "Six random samples from Model 2's parameters' posterior distributions, and their summaries.",
    escape = FALSE,
    placement = "h"
  )
```

```{r}
#| label: calc-2-heterogeneity

# Calculate all heterogeneity metrics and differences
post2 <- post2 |> 
  mutate(
    # Persons
    dist_person_free = dist_normal(b_taskFree, `sd_person__Intercept:taskFree`),
    dist_person_cued = dist_normal(b_taskCued, `sd_person__Intercept:taskCued`),
    pd_person_free = 1 - cdf(dist_person_free, 0),
    pd_person_cued = 1 - cdf(dist_person_cued, 0),
    ratio_person_free = abs(`sd_person__Intercept:taskFree` / b_taskFree),
    ratio_person_cued = abs(`sd_person__Intercept:taskCued` / b_taskCued),
    hi_low_free = b_taskFree + qnorm(0.05) * `sd_person__Intercept:taskFree`,
    hi_high_free = b_taskFree + qnorm(0.95) * `sd_person__Intercept:taskFree`,
    hi_low_cued = b_taskCued + qnorm(0.05) * `sd_person__Intercept:taskCued`,
    hi_high_cued = b_taskCued + qnorm(0.95) * `sd_person__Intercept:taskCued`,
    # Differences
    b_difference = b_taskCued - b_taskFree,
    sd_person_difference = 
      `sd_person__Intercept:taskCued` - `sd_person__Intercept:taskFree`,
    sd_person_ratio = 
      `sd_person__Intercept:taskCued` / `sd_person__Intercept:taskFree`,
    pd_person_difference = pd_person_cued - pd_person_free,
    ratio_person_difference = ratio_person_cued - ratio_person_free,
    hi_low_difference = hi_low_cued - hi_low_free,
    hi_high_difference = hi_high_cued - hi_high_free,
    # Targets
    dist_target_free = dist_normal(b_taskFree, `sd_target__taskFree`),
    dist_target_cued = dist_normal(b_taskCued, `sd_target__taskCued`),
    pd_target_free = 1 - cdf(dist_target_free, 0),
    pd_target_cued = 1 - cdf(dist_target_cued, 0),
    ratio_target_free = abs(`sd_target__taskFree` / b_taskFree),
    ratio_target_cued = abs(`sd_target__taskCued` / b_taskCued),
    # Differences
    sd_target_difference = 
      `sd_target__taskCued` - `sd_target__taskFree`,
    pd_target_difference = pd_target_cued - pd_target_free,
    ratio_target_difference = ratio_target_cued - ratio_target_free,
    .before = starts_with("cor_")
  ) |> 
  select(-starts_with("dist_"))

post2_sum <- post2 |> 
  mutate(
    across(b_taskFree:b_taskCued, list(p = ~pnorm(.x)))
  ) |> 
  summarize_draws(
    mean, ~quantile2(.x, probs = c(0.025, 0.975))
  ) |> 
  mutate(
    x = str_glue("{number(mean, .01)} [{number(q2.5, .01)}, {number(q97.5, .01)}]")
  )
```

## Comparing between-person heterogeneity across tasks

Descriptively, we reproduced @mahVariabilitySubjectsFree2023's finding that participants' memory performance was more heterogeneous in the cued recall than in the free recall experiment (columns 4 and 5 in Table \@ref(tab:tbl-samples-2)). We show the relevant estimated quantities, and the implied heterogeneity distributions in Figure \@ref(fig:fig-2-person).

The top panel of Figure \@ref(fig:fig-2-person)A illustrates the posterior distributions of memory performance for the average person in the free and cued recall tasks, and their difference (cued - free recall). Notice that the model's parameters refer to probits ("z-scores"): While recall performance was `r post2_sum$x[post2_sum$variable=="b_taskFree"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued"]` probits in the free and cued recall conditions, respectively, the corresponding probabilities were `r post2_sum$x[post2_sum$variable=="b_taskFree_p"]` and `r post2_sum$x[post2_sum$variable=="b_taskCued_p"]`. However, this conversion does not work in general and we therefore restrict descriptions to probits in what follows.

```{r}
#| label: fig-2-person-calc

x <- seq(-2.2, 2.2, length = 201)

tmp_person <- bind_rows(
  "Free" = post2 |>
    select(b_taskFree, `sd_person__Intercept:taskFree`) |>
    calculate_density(b_taskFree, `sd_person__Intercept:taskFree`) |> 
    select(x, pdf, cdf),
  "Cued" = post2 |>
    select(b_taskCued, `sd_person__Intercept:taskCued`) |>
    calculate_density(b_taskCued, `sd_person__Intercept:taskCued`) |> 
    select(x, pdf, cdf),
  .id = "Task"
)  |> 
  group_by(x, Task) |>
  curve_interval(.width = 0.9)
```

```{r}
#| label: model-2-ratio
#| include: false

# Some ratios are weird
post2 |> 
  select(
    starts_with("b_"), 
    starts_with("sd_person"), 
    starts_with("ratio_person"), 
    -ends_with("_difference")
  ) |> 
  select(
    contains("free")
  ) |> 
  filter(ratio_person_free > 10)
```

```{r}
#| label: fig-2-person
#| fig.height: 3.2
#| fig.env: "figure*"
#| fig.cap: "Estimated between-person heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2. A. Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs. Differences calculated as Cued - Free recall. B. Probability density (top) and cumulative density functions (bottom) of the two groups' heterogeneity distributions (green: free recall, red: cued recall). The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance."

p2_pars_person <- post2 |> 
  select(contains("b_") | contains("person"), -sd_person_ratio) |> 
  pivot_longer(everything()) |> 
  mutate(
    name = str_to_lower(name) |> 
      str_remove("_person") |> 
      str_remove("_intercept:task") |> 
      str_remove("task")
  ) |> 
  separate(name, c("type", "group")) |> 
  mutate(
    type = factor(
      type, 
      levels = c("b", "sd", "pd", "ratio"),
      labels = c("Mean", "SD", "Pr(Above chance)", "Ratio")
    ),
    group = factor(
      group,
      levels = c("free", "cued", "difference"),
      labels = c("Free", "Cued", "Difference")
    ) |> fct_rev(),
    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
  ) |> 
  ggplot(aes(value, group, col = group, fill = group)) +
  scale_color_brewer(
    "Recall task",
    palette = "Dark2",
    aesthetics = c("color", "fill"),
    direction = -1
  ) +
  scale_y_discrete(
    expand = expansion(c(0.1, 0.25))
  ) +
  scale_x_continuous(
    "Parameter value",
    expand = expansion(0.025)
  ) +
  geom_vline(
    xintercept = 0, linewidth = .2, alpha = .25
  ) +
  stat_histinterval(
    slab_alpha = .25,
    breaks = 50,
    normalize = "panels",
    .width = .95,
    size = 1,
    justification = 0.01
  ) +
  facet_wrap("type", ncol = 1, scales = "free") +
  theme(
    legend.position = "none",
    axis.title.y = element_blank()
  )

p2_cdf_person <- tmp_person |> 
  ggplot(aes(x = x)) +
  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
  geom_vline(
    xintercept = 0, linewidth = .2, alpha = .25
  ) +
  scale_color_brewer(
    "Recall task",
    palette = "Dark2",
    aesthetics = c("color", "fill"),
    direction = -1
  ) +
  scale_x_continuous(
    TeX("$\\Phi(Pr(Correct))$"),
    expand = expansion(0.01),
    breaks = extended_breaks(5),
    sec.axis = sec_axis(
      name = "Proportion correct",
      trans = ~pnorm(.), 
      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
    )
  ) +
  scale_y_continuous(
    breaks = extended_breaks(7),
    expand = expansion(c(0.005, 0.005))
  ) +
  ylab("CDF") +
  geom_ribbon(
    alpha = .2, col = NA
  ) +
  geom_line(
    linewidth = .5
  ) +
  theme(
    legend.position = "none",
    panel.border = element_rect(fill = NA)
  )

p2_pdf_person <- p2_cdf_person %+% 
  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
  ylab("PDF")

p2_cdf_person_inset <- post2 |> 
  select(pd_person_free, pd_person_cued) |> 
  pivot_longer(everything()) |> 
  ggplot() +
  scale_y_01 +
  scale_x_continuous(
    expand = expansion(0)
  ) +
  scale_color_brewer(
    "Recall task",
    palette = "Dark2",
    aesthetics = c("color", "fill"),
    direction = -1
  ) +
  stat_slab(
    aes(y = value, fill = name),
    side = "right",
    .width = .95,
    size = 1,
    scale = 0.5,
    alpha = .2,
    justification = 0.025
  ) +
  stat_pointinterval(
    aes(y = value, col = name),
    side = "right",
    .width = .95,
    size = 1,
    justification = 0.025
  ) +
  theme_void() +
  theme(legend.position = "none")

p2_cdf_person <- p2_cdf_person +
  inset_element(
    p2_cdf_person_inset,
    0, 0, .2, 1, 
    align_to = "panel", 
    on_top = TRUE
  )

p2_pcdf_person <- p2_pdf_person / p2_cdf_person

(p2_pars_person | wrap_elements(full = p2_pcdf_person)) +
  plot_annotation(tag_levels = "A")
```

Second, and more importantly, the second row in Figure \@ref(fig:fig-2-person)A describes the posterior distributions of the between-person standard deviations in memory ability in the free and cued recall tasks, and their difference (cued - free recall). The standard deviation was `r post2_sum$x[post2_sum$variable=="sd_person_difference"]` probits greater in the cued recall task (ratio: `r post2_sum$x[post2_sum$variable=="sd_person_ratio"]`).

The third row of Figure \@ref(fig:fig-2-person)A shows the estimated proportions of individuals whose memory performance is above chance in the two tasks ($p^+$), and their difference. The model estimates the proportion of individuals who perform above chance to be `r post2_sum$x[post2_sum$variable=="pd_person_difference"]` greater in the cued than in the free recall task. Notice that the this quantity refers to population proportions and is not a z-score.

Perhaps surprisingly, even though the absolute measures of heterogeneity differed greatly between the two recall tasks, the bottom panel of Figure \@ref(fig:fig-2-person)A shows that the degree of relative heterogeneity is virtually identical across the two tasks. (We truncated the Ratio axis at [-10, 10] because estimates of the ratio are extreme for estimated distributions with very small average effects but large standard deviations. The difference in ratios is very uncertain partly for the same reason.)

We also depict the heterogeneity distributions as PDFs and CDFs in Figure \@ref(fig:fig-2-person)B. These figures, especially the bottom panel, allow for a concise description of (differences in) heterogeneity in the two tasks: First, we see that the majority of the free recall CDF (green) is to the left of zero (chance), indicating that the majority of this population is predicted to perform worse than chance. This information is described in more detail in the small posterior densities and point-intervals on the left y-axis: The model predicts above-chance performance only for a proportion of `r post2_sum$x[post2_sum$variable=="pd_person_free"]` of the population. Second, we see that the slope of the cued recall CDF (red) is less steep than that of the free recall CDF: The between-person distribution of memory abilities is more dispersed in the cued than in the free recall task.

Finally, we turn to the heterogeneity interval (HI). The $HI_{90}$'s lower bound in the free recall task is `r post2_sum$x[post2_sum$variable=="hi_low_free"]`, and `r post2_sum$x[post2_sum$variable=="hi_low_cued"]` in the cued recall task. While this 5th percentile of the heterogeneity distribution was not credibly different across the two tasks (Cued - Free recall; `r post2_sum$x[post2_sum$variable=="hi_low_difference"]`), the 95th percentiles differed at the 95% credibility level (the Cued recall upper $HI_90$ limit was `r post2_sum$x[post2_sum$variable=="hi_high_difference"]` probits greater). Studying Figure \@ref(fig:fig-2-person)B closely makes another implication of the different standard deviations clear: While the average person likely has greater memory performance in the cued recall task, the model predicts that there are also more individuals with very poor performances in the cued recall condition.

## Comparing between-target heterogeneity across tasks

Between-person heterogeneity is typically the more theoretically important phenomenon for psychologists than differences in model parameters between other randomly sampled study units, such as stimuli. However, examining heterogeneity in other sampled units can be both theoretically and methodologically important. Recognizing this difference, and for the sake of our example analyses we next turn to our second and third questions regarding potential differences and consistencies in between-target word heterogeneity.

```{r}
#| label: fig-2-target-calc

x <- seq(-2.2, 2.2, length = 201)

tmp_target <- bind_rows(
  "Free" = post2 |>
    select(b_taskFree, sd_target__taskFree) |>
    calculate_density(b_taskFree, `sd_target__taskFree`) |> 
    select(x, pdf, cdf),
  "Cued" = post2 |>
    select(b_taskCued, `sd_target__taskCued`) |>
    calculate_density(b_taskCued, `sd_target__taskCued`) |> 
    select(x, pdf, cdf),
  .id = "Task"
)  |> 
  group_by(x, Task) |>
  curve_interval(.width = 0.9)
```

```{r}
#| label: fig-2-target
#| fig.height: 4
#| fig.width: 7.4
#| fig.env: "figure*"
#| fig.cap: "Estimated between-target word heterogeneity in memory performance in Free recall and Cued recall tasks from Model 2. A. Histograms of 4,000 posterior draws from the model parameters and their transformations, with points and intervals showing posterior means and 95\\%CIs. Differences calculated as Cued - Free recall. B. Probability density (top) and cumulative density functions (bottom) of the two tasks' heterogeneity distributions (green: free recall, red: cued recall). The densities, points, and intervals on the left y-axis of the bottom panel indicate approximate posterior densities, with means and 95\\%CIs, of the proportions of the populations with memory performance above chance. C. Posterior mean (dark), and 100 posterior draws (light) of the correlation between target words' proportions correct in the free (x-axis) and cued recall (y-axis) tasks. Ellipses indicate the 90th percentile of the bivariate normal distribution."

p2_pars_target <- post2 |> 
  select(contains("target")) |> 
  rename_with(
    ~str_to_lower(.) |> 
      str_remove("target_") |> 
      str_remove_all("_task")
  ) |> 
  pivot_longer(everything()) |> 
  separate(name, c("type", "group"), extra = "merge") |> 
  mutate(
    type = factor(
      type, 
      levels = c("sd", "pd", "ratio", "cor"),
      labels = c("SD", "Pr(Above chance)", "Ratio", "Cor(Free, Cued)")
    ),
    group = factor(
      group,
      levels = c("free", "cued", "difference", "free_cued"),
      labels = c("Free", "Cued", "Difference", "Correlation")
    ) |> fct_rev(),
    value = if_else(type == "Ratio" & abs(value) > 10, NaN, value)
  ) |> 
  ggplot(aes(value, group, color = group, fill = group)) +
  scale_color_manual(
    "Recall task",
    values = c(rep("#7570B3", 2), "#D95F02", "#1B9E77"),
    aesthetics = c("color", "fill")
  ) +
  scale_y_discrete(
    expand = expansion(c(0.1, 0.25))
  ) +
  scale_x_continuous(
    "Parameter value",
    expand = expansion(0.1)
  ) +
  geom_vline(
    xintercept = 0, linewidth = .2, alpha = .25
  ) +
  stat_histinterval(
    breaks = 50,
    normalize = "panels",
    .width = .95,
    size = 1,
    slab_alpha = 0.2,
    justification = 0.01
  ) +
  facet_wrap("type", ncol = 1, scales = "free") +
  theme(
    legend.position = "none",
    axis.title.y = element_blank()
  )

p2_cdf_target <- tmp_target |> 
  ggplot(aes(x = x)) +
  aes(y = cdf, ymin = cdf.lower, ymax = cdf.upper, col = Task, fill = Task) +
  geom_vline(
    xintercept = 0, linewidth = .2, alpha = .25
  ) +
  scale_color_brewer(
    "Recall task",
    palette = "Dark2",
    aesthetics = c("color", "fill"),
    direction = -1
  ) +
  scale_x_continuous(
    TeX("$\\Phi(Pr(Correct))$"),
    expand = expansion(0.01),
    breaks = extended_breaks(5),
    sec.axis = sec_axis(
      name = "Proportion correct",
      trans = ~pnorm(.), 
      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
    )
  ) +
  scale_y_continuous(
    breaks = extended_breaks(7),
    expand = expansion(c(0.005, 0.005))
  ) +
  ylab("CDF") +
  geom_ribbon(
    alpha = .2, col = NA
  ) +
  geom_line(
    linewidth = .5
  ) +
  theme(
    legend.position = "none",
    panel.border = element_rect(fill = NA)
  )

p2_pdf_target <- p2_cdf_target %+% 
  aes(y = pdf, ymin = pdf.lower, ymax = pdf.upper) +
  ylab("PDF")

p2_cdf_target_inset <- post2 |> 
  select(pd_target_free, pd_target_cued) |> 
  pivot_longer(everything()) |> 
  ggplot() +
  scale_y_01 +
  scale_x_continuous(
    expand = expansion(0)
  ) +
  scale_color_brewer(
    "Recall task",
    palette = "Dark2",
    aesthetics = c("color", "fill"),
    direction = -1
  ) +
  stat_slab(
    aes(y = value, fill = name),
    side = "right",
    .width = .95,
    size = 1,
    scale = 0.5,
    alpha = .2,
    justification = 0.025
  ) +
  stat_pointinterval(
    aes(y = value, col = name),
    side = "right",
    .width = .95,
    size = 1,
    justification = 0.025
  ) +
  theme_void() +
  theme(legend.position = "none")

p2_cdf_target <- p2_cdf_target +
  inset_element(
    p2_cdf_target_inset,
    0, 0, .2, 1, 
    align_to = "panel", 
    on_top = TRUE
  )

p2_pcdf_target <-  p2_pdf_target / p2_cdf_target

# Ellipses
.ellipse <- function(data) {
  data |>
    mutate(
      e = pmap(
        list(
          cor_target__taskFree__taskCued,
          sd_target__taskFree,
          sd_target__taskCued,
          b_taskFree,
          b_taskCued
        ),
        ~ ellipse(
          x = ..1,
          scale = c(..2, ..3),
          centre = c(..4, ..5),
          level = .90
        ) |>
          as_tibble()
      )
    ) |>
    unnest(e)
}
samples <- as_draws_df(
  fit2,
  variable = c(
    "b_",
    "sd_",
    "cor_"
  ),
  regex = TRUE
)

set.seed(99)
ellipse_samples <- samples |>
  slice_sample(n = 100) |>
  .ellipse()

ellipse_median <- samples |>
  summarise_draws(median) |>
  pivot_wider(names_from = variable, values_from = median) |>
  .ellipse()

# Target word coefficients, data and model
dat2_target_coefs <- left_join(
  dat |> 
    summarise(
      p = qnorm(mean(accuracy)),
      .by = c(task, target)
    ) |> 
    pivot_wider(
      names_from = task, values_from = p
    ),
  coef(fit2)$target |> 
    as.data.frame() |> 
    rownames_to_column("target") |> 
    tibble() |> 
    select(
      target, 
      .Free = Estimate.taskFree,
      .Cued = Estimate.taskCued
    )
)

p2_ellipse <- ellipse_median |>
  ggplot(aes(x, y)) +
  coord_cartesian(
    xlim = c(-1.8, 1.5),
    ylim = c(-1.5, 1.8)
  ) +
  geom_vline(
    xintercept = 0, linewidth = .1, linetype = "dashed"
  ) +
  geom_hline(
    yintercept = 0, linewidth = .1, linetype = "dashed"
  ) +
  geom_abline(
    linewidth = .1, linetype = "dashed"
  ) +
  scale_x_continuous(
    TeX("$\\Phi(Pr(Correct))~Free~recall$"),
    breaks = seq(-1.5, 1.5, by = .5),
    sec.axis = sec_axis(
      trans = ~ pnorm(.),
      name = "Pr(Correct) Free recall",
      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
    )
  ) +
  scale_y_continuous(
    TeX("$\\Phi(Pr(Correct))~Cued~recall$"),
    breaks = seq(-1.5, 1.5, by = .5),
    sec.axis = sec_axis(
      trans = ~ pnorm(.),
      name = "Pr(Correct) Cued recall",
      breaks = c(0.05, .1, .3, .5, .7, .9, .95),
      labels = c(".05", ".1", ".3", ".5", ".7", ".9", ".95")
    )
  ) +
  geom_path(
    data = ellipse_samples,
    aes(group = .draw),
    alpha = .075, color = "#7570B3"
  ) +
  geom_path(
    color = "#7570B3"
  ) +
  # geom_segment(
  #   data = dat2_target_coefs,
  #   aes(x = Free, xend = .Free, y = Cued, yend = .Cued)
  # ) +
  # geom_point(
  #   data = dat2_target_coefs,
  #   aes(x = Free, y = Cued),
  #   shape = 21, fill = "white"
  # ) +
  # geom_point(
  #   data = dat2_target_coefs,
  #   aes(x = .Free, y = .Cued),
  #   color = "dodgerblue4"
  # ) +
  theme(
    aspect.ratio = 1
  )

(wrap_elements(full = p2_pars_target) | wrap_elements(full = p2_pcdf_target) | p2_ellipse) +
  plot_annotation(tag_levels = "A")
```

The results regarding differences in memory performance heterogeneity across target words were similar to those as observed regarding heterogeneity in people's memory performances. Figure \@ref(fig:fig-2-target)A shows that heterogeneity (standard deviations) in memory performance was greater when words appeared in the cued recall task. The interpretation of this difference is quantitatively similar to that observed about people above: Both people and target words exhibit greater memory task performance variability in the cued recall than in the free recall task. However, because @mahVariabilitySubjectsFree2023 used the same target words across the two tasks, this interpretation is subtly more complex: This difference holds even when the exact same units---target words, in this example---are used in the two different tasks.

Moreover, we observe similar differences in between-target word heterogeneity between the two tasks as we did above regarding between-person heterogeneity: The model predicts a greater proportion of words to elicit greater than chance performance in the cued recall than in the free recall task. Yet, the ratio of the heterogeneity distribution's standard deviation to its mean again appeared nearly identical across the two tasks.

This analysis of target word heterogeneity afforded an additional piece of information, because the same target words were used across the two tasks, and thus allowed answering our third question. There was a clear positive correlation between target words' rates of correct responses across the free and cued recall tasks (bottom panel of Figure \@ref(fig:fig-2-target)A, and Figure \@ref(fig:fig-2-target)C). The posterior mean and 95%CI of this correlation was `r post2_sum$x[post2_sum$variable=="cor_target__taskFree__taskCued"]`. What this correlation means substantively is that words that are likely better recalled in the free recall task are also likely to be those that are better recalled in the cued recall task. @bolgerCausalProcessesPsychology2019 found a conceptually similar result regarding valence effects' stability across time but within individuals: Individuals whose valence effect was stronger at Time 1 were also those whose valence effect was likely to be stronger at Time 2, one week later.

Although outside our current scope, we note that our study of Model 2's results might indicate exciting new avenues for this line of inquiry. One explanation for the between-task difference in between-person heterogeneity is that participants might adopt different strategies for remembering in the two different recall tasks [@mahVariabilitySubjectsFree2023]. Our additional results suggest that such a mechanism may not be a complete account of differences in memory performance heterogeneity: Target words are presumably invariant regarding memory strategies, yet we find that recall accuracy is more heterogeneous across target words in the cued recall task than in the free recall task (Figure \@ref(fig:fig-2-target)A). Moreover, we observed across two kinds of study units (people, target words) that the ratio of the between-unit standard deviation to the average effect was nearly identical across the free and cued recall tasks.

# Discussion

In the current work, we illustrated the use of practical descriptors of heterogeneity in psychological phenomena with examples drawn from social and cognitive psychology. Our aim was to incrementally build on the work of @bolgerCausalProcessesPsychology2019 and others---who have described the importance and available methods for examining heterogeneity in causal effects in psychology---by describing how it is both critically important and practically feasible to incorporate uncertainty in analyses and descriptions of heterogeneity. We believe our currently proposed methods to be particularly useful to researchers because they allow doing just that: Incorporating uncertainty in the analyses at every stage, from modelling to communicating resulting inferences.

We emphasized throughout that a probabilistic ("bayesian") approach is uniquely positioned to answer the needs of researchers interested in heterogeneity. Bayesian methods allow carrying uncertainty forward from model parameters to descriptors of heterogeneity and beyond. The resulting metrics are useful because they not only convey analysts' expectations regarding heterogeneity, but more fully convey their states of knowledge regarding heterogeneity, including degrees of certainty. In addition, probabilistic modelling, by returning a matrix of samples from the posterior distribution, enables practically straightforward solutions whereby analysts can use familiar data wrangling techniques to easily compare various heterogeneity descriptors across groups. We further elucidate these points in our Online Supplementary Analysis [todo].

We believe that psychology, broadly speaking, is methodologically and theoretically at a ripe position to taking a meta-theoretic step forward and try to incorporate an understanding of heterogeneity into substantive theories [@bolgerCausalProcessesPsychology2019]. In order to do so, descriptions of heterogeneity must include measures of uncertainty, and we hope the techniques illustrated here help researchers do so.

## Limitations

In our example analyses, we have brushed many important modelling decisions under the rug in order to focus on the main topic of heterogeneity. First, in example 1 we analyzed reaction times by simply log-transforming reaction times. More informative analyses of RTs would make use of models that make more realistic assumptions about the data generating process underlying reaction time responses, but here we necessarily excluded this complication for reasons of brevity.

Our exposition and interpretation of heterogeneity relied on a critical assumption in line with standard practices in multilevel and generalized linear mixed modelling; that of (multivariate) normality of the group-level parameters. Assuming that random effects are normally distributed is a computationally and conceptually useful fiction, and we recognize that it is unlikely to hold exactly in real psychological phenomena. Haaf, Rouder, and colleagues have explored alternatives to continuous normal distributions of random effects [e.g., @haafDevelopingConstraintBayesian2017; @haafDonAccountingVariability2019].

## Conclusion

We hope that the conceptual, computational, and graphical tools that we have discussed here prove useful to researchers interested in better understanding heterogeneity in psychological phenomena.

# Data and code availability

All the code supporting this manuscript is available at <https://github.com/mvuorre/heterogeneity-uncertainty> and archived on Zenodo at [todo]. We reused openly available data from @bolgerCausalProcessesPsychology2019 and @mahVariabilitySubjectsFree2023.

# Author contributions

<!-- https://casrai.org/credit/ -->

Conceptualization: MV, NB\
Formal Analysis: MV\
Methodology: MV, NB, MK\
Project Administration: MV\
Software: MV, MK\
Visualization: MV, MK, NB\
Writing – Original Draft: MV\
Writing – Review & Editing: MV, NB, MK

# Competing interests

The author(s) declare no competing interests.

# References {-}

::: {#refs custom-style="Bibliography"}
:::

<!-- Prepare appendix section -->
\clearpage
\setcounter{page}{1}
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}
\onecolumn

# Appendix

```{r}
#| label: intervals
#| fig.height: 2.4
#| fig.cap: Bayesian estimates of the 90% heterogeneity interval of valence effects. A. Scatterplot of 4,000 posterior samples of the lower (x-axis) and upper (y-axis) limits of the $HI_{90}$. B. Histograms of 4,000 samples of the HI90 lower (left) and upper (right) limits with their posterior means and 95%CIs as points and intervals. C. 100 random samples from the $HI_{90}$ posterior distribution, with the posterior mean heterogeneity interval superimposed in a darker shade of blue.

p1 <- post1 |>
  mutate(
    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
  ) |>
  ggplot(aes(hi_low, hi_high)) +
  labs(x = "HI90 lower limit", y = "HI90 upper limit") +
  geom_point(
    alpha = .25,
    shape = 1,
    size = 0.75,
    col = "dodgerblue1"
  ) +
  theme(aspect.ratio = 1)

p2 <- post1 |>
  mutate(
    hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
    hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
  ) |>
  select(Sample, hi_low, hi_high) |>
  pivot_longer(-Sample) |>
  ggplot(aes(value, group = name)) +
  scale_y_continuous(
    expand = expansion(c(0, 0.1))
  ) +
  scale_x_continuous(
    "Valence effects"
  ) +
  stat_histinterval(
    breaks = 100,
    fill = alpha("dodgerblue1", .33),
    color = "dodgerblue4",
    .width = c(.95),
    justification = .01
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p3 <- distinct(post1_density_samples, Sample, b_valence1, sd_person__valence1) |>
  arrange(b_valence1) |>
  mutate(i = 1:n()) |>
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1)
  ) |>
  ggplot() +
  scale_x_continuous(
    "Valence effect",
    expand = expansion(0.025),
    breaks = extended_breaks()
  ) +
  scale_y_continuous(
    expand = expansion(0.025)
  ) +
  labs(y = "Posterior sample") +
  stat_pointinterval(
    aes(xdist = dist, y = i),
    .width = .9,
    interval_size_range = c(.15, .3),
    alpha = .5,
    color = "dodgerblue1"
  ) +
  geom_pointinterval(
    data = post1 |>
      summarise(
        across(c(b_valence1, sd_person__valence1), mean)
      ) |>
      mutate(
        hi_low = b_valence1 + qnorm(0.05) * sd_person__valence1,
        hi_high = b_valence1 + qnorm(0.95) * sd_person__valence1
      ),
    aes(y = 50, x = b_valence1, xmin = hi_low, xmax = hi_high),
    color = "dodgerblue4"
  )

(p1 | p2 | p3) +
  # coord_cartesian(xlim = c(-0.6, 0.3)) &
  # scale_x_continuous(
  #   "Valence effect",
  #   expand = expansion(0),
  #   breaks = extended_breaks(7)
  # )) +
  # plot_layout(widths = c(6, 4)) +
  plot_annotation(tag_levels = "A")
```